soa

### 2024-08-28 inst/manuscript_figures.Rmd
vignette ds changes

f <- file.path(pca_dir, "pca_grid_2000")
ggsave(paste0(f, ".png"), p, width=20, height=15)
ggsave(paste0(f, ".pdf"), p, width=20, height=15)

@@

### Pearson correlates
```{r}
# Example data (replace with your actual data)
methods <- c("none_q3", "none_tmm", "none_rle", "none_qn", 
             "bgsub_q3", "bgsub_tmm", "bgsub_rle", "bgsub_qn", 
             "qq_q3", "qq_tmm", "qq_rle", "qq_qn")

@@

### PC1, PC2, PC3, PC4, PC5 for qq_qn
``` {r pcs}
s <- fds_qq_qn$samples
pcs <- c("PC1", "PC2", "PC3", "PC4", "PC5")
pca_tbl <- pca_res_qq_qn$x
tbl <- as_tibble(pca_tbl[,pcs], rownames="aoi")

@@

f <- file.path(pca_dir, "pca_12345_grid")
ggsave(paste0(f, ".png"), p, width=25, height=20)
ggsave(paste0(f, ".pdf"), p, width=25, height=20)
```

@@

  p <- ggplot(res, aes(x=log2fc, y=-log10(padj), color=de, size=de)) +
    geom_point(alpha=0.7) +
    geom_text_repel(data=label_data, color="black", size=3, aes(label=gene), max.overlaps=Inf) +
    scale_color_manual(values=c("no"="grey", "dn"="blue", "up"="red")) +
    scale_size_manual(values=c("no"=1, "dn"=2, "up"=2)) +
    theme_minimal() +
    theme(legend.position="bottom") +
    theme(axis.line = element_line(color = "black")) +
    labs(x="log2fc", y="-log10(padj)")
  return(p)
}

@@

# heatmap functions
plot_pheatmap <- function(m, 
                          annot_col, 
                          annot_row, 
                          annot_colors,
                          filename, width, height,
                          xmin=-6, xmax=6) {
  pheatmap(m,
           scale = "none",
           color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
           breaks = seq(xmin, xmax, length.out=51),
           annotation_col = annot_col,
           annotation_row = annot_row,
           annotation_colors = annot_colors,
           border_color = NA,
           cluster_rows = TRUE,
           cluster_cols = TRUE,
           clustering_distance_rows = "euclidean",
           clustering_distance_cols = "euclidean",
           clustering_method = "ward.D2",
#           cellwidth = 6,
#           cellheight = 6,
           fontsize = 6,
           fontsize_row = 6,
           fontsize_col = 6,
           filename = filename, 
           width = width,
           height = height)
}

# setup heatmap annotations
gene_annot <- fds_qq_qn$meta 
# %>%
#   left_join(hpa_pdac, by=c("gene"="gene_symbol")) %>%
#   left_join(cptac_de_merged, by=c("gene"="gene_symbol")) %>%
#   mutate(de_cptac_rna = replace_na(de_cptac_rna, "na"),
#          de_cptac_prot = replace_na(de_cptac_prot, "na"))
annot_row <- gene_annot %>%
  select(gene) %>%
  column_to_rownames("gene") %>%
  as.data.frame()
annot_col <- column_to_rownames(fds_qq_qn$samples, "aoi") %>%
  select(patient, study, cluster_name) %>%
  as.data.frame()

@@

## FIGURE 6: GSEA
``` {r gsea}

>>>>

f <- file.path(pca_dir, "pca_grid_500")
ggsave(paste0(f, ".png"), p, width=20, height=15)
ggsave(paste0(f, ".pdf"), p, width=20, height=15)

@@

### Pearson correlates
```{r}

methods <- c("none_q3", "none_tmm", "none_rle", "none_qn", 
             "bgsub_q3", "bgsub_tmm", "bgsub_rle", "bgsub_qn", 
             "qq_q3", "qq_tmm", "qq_rle", "qq_qn")

@@

### PC1, PC2, PC3, PC4, PC5 for qq_qn
``` {r pcs}
s <- fds_qq_qn$samples
pcs <- c("PC1", "PC2", "PC3", "PC4")
pca_tbl <- pca_res_qq_qn$x
tbl <- as_tibble(pca_tbl[,pcs], rownames="aoi")

@@

f <- file.path(pca_dir, "pca_1234_grid")
ggsave(paste0(f, ".png"), p, width=25, height=20)
ggsave(paste0(f, ".pdf"), p, width=25, height=20)
```

@@

  p <- ggplot(res, aes(x=log2fc, y=-log10(padj + 1e-300), color=de, size=de)) +
    geom_point(alpha=0.7) +
    geom_text_repel(data=label_data, color="black", size=3, aes(label=gene), max.overlaps=Inf) +
    scale_color_manual(values=c("no"="grey", "dn"="blue", "up"="red")) +
    scale_size_manual(values=c("no"=1, "dn"=2, "up"=2)) +
    theme_minimal() +
    theme(legend.position="bottom") +
    theme(axis.line = element_line(color = "black")) +
    labs(x="log2fc", y="-log10(padj)")
  return(p)
}

@@

# heatmap functions
plot_pheatmap <- function(m, 
                          annot_col,
                          # annot_row, 
                          annot_colors,
                          filename, width, height,
                          xmin=-6, xmax=6) {
  pheatmap(m,
           scale = "none",
           color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
           breaks = seq(xmin, xmax, length.out=51),
           annotation_col = annot_col,
           # annotation_row = annot_row,
           # annotation_colors = annot_colors,
           border_color = NA,
           cluster_rows = TRUE,
           cluster_cols = TRUE,
           clustering_distance_rows = "euclidean",
           clustering_distance_cols = "euclidean",
           clustering_method = "ward.D2",
#           cellwidth = 6,
#           cellheight = 6,
           fontsize = 6,
           fontsize_row = 6,
           fontsize_col = 6,
           filename = filename, 
           width = width,
           height = height)
}

# setup heatmap annotations
# column annotations: show clusters, study
# no row annotations
gene_annot <- fds_qq_qn$meta
# %>%
#   left_join(hpa_pdac, by=c("gene"="gene_symbol")) %>%
#   left_join(cptac_de_merged, by=c("gene"="gene_symbol")) %>%
#   mutate(de_cptac_rna = replace_na(de_cptac_rna, "na"),
#          de_cptac_prot = replace_na(de_cptac_prot, "na"))
# annot_row <- gene_annot %>%
#   select(gene) %>%
#   column_to_rownames("gene") %>%
#   as.data.frame()
annot_col <- column_to_rownames(fds_qq_qn$samples, "aoi") %>%
  select(study, cluster_name) %>%
  as.data.frame()

@@

## FIGURE 6: GSEA
``` {r gsea}
# read msigdb data
f <- file.path(data_dir, "msigdb.v2023.2.Hs.json")
msigdb_gene_sets <- read_msigdb_json_file(f)
# determine gene universe
msigdb_genes <- intersect(fds$meta$gene, msigdb_gene_sets$gene_symbol)
# filter genes not in our dataset
msigdb_gene_sets <- msigdb_gene_sets %>%
  dplyr::filter(gene_symbol %in% fds$meta$gene) %>%
  distinct(gs_cat, gs_name, gene_symbol, .keep_all = TRUE)

# gene set categories
gs_hallmark <- dplyr::filter(msigdb_gene_sets, gs_cat == "H")
gs_hallmark <- split(x = gs_hallmark$gene_symbol, f = gs_hallmark$gs_name)

# pass log fold change into gsea with hallmark gene sets (sending new code)

get_de_ranks <- function(xde, mycontrast) {
  ranks <- xde %>% 
    dplyr::filter(contrast == mycontrast) %>%
    select(gene, log2fc, padj) %>%
    mutate(rank = log2fc)
  ranks = sort(setNames(ranks$rank, ranks$gene), decreasing = TRUE)
  return(ranks)
}

plot_gsea_enrichment <- function(x, ranks, a, gs) {
  txt_stat <- paste0("ES=", round(x$ES,2), " NES=", round(x$NES, 2), " padj=", format(x$padj, scientific=TRUE, digits=3))
  txt_title <- paste0("ranks: ", a, " gs: ", x$pathway)
  p <- plotEnrichment(gs[[x$pathway]], ranks) +
    annotate(geom="text", x=150, y=0.1, label=txt_stat, hjust=0) +
    labs(title = txt_title, 
         xlab = "Rank",
         ylab = "Enrichment Score") +
    theme(plot.title = element_text(size=8))
  return(p)
}

# fgsea_run_batch <- function(xde, xgs, 
#                             fgsea_eps=0,
#                             fgsea_nperms=10000, 
#                             fgsea_nproc=0) {
#   fgsea_scoretypes <- c("pos", "neg")
#   gsea <- NULL
#   for (xcontrast in unique(xde$contrast)) {
#     ranks <- get_de_ranks(xde, xcontrast)
#     for (fgsea_scoretype in fgsea_scoretypes) {
      res <- fgsea(pathways=xgs, 
                   stats=ranks, 
                   eps=fgsea_eps, 
                   scoreType = fgsea_scoretype,
                   nPermSimple=fgsea_nperms, 
                   nproc=fgsea_nproc)
      res <- mutate(res, 
                    contrast = xcontrast, 
                    dir = fgsea_scoretype)
      gsea <- bind_rows(gsea, res)      
#     }
#   }
#   return(gsea)
# }

#
# Run full analysis
#
gsea <- NULL

# hallmark
res <- fgsea_run_batch(de, gs_hallmark, 
                       fgsea_eps=0, fgsea_nperms=gsea_nperms, fgsea_nproc=10)
gsea <- bind_rows(gsea, res)
gsea_all <- gsea
```
